The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-05-30 brains-introduction.tex
chore/todos: update some todos from Gureckis' edited version

@@

>>>>

\todo{
\#14: Rework to
(1) present shifts in biology (eco-evo-devo) rejoining attempts to bridge social/cognitive rift and study knowledge transmission (cultural evolution),
(2) present current problem of testing theories to incorporate those discoveries,
(3) present big data literature which can bring measurement tools,
(4) introduce the question of cultural attractors,
(5) to be studied through quotes on the web: it's the proposal of an online approach, we'll be describing and cross-checking with existing literature (vs. predicting, since there are too many possible things to predict)
}

\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}

@@

\todo{
\#15: add
(1) sentence recall (Potter \& Lombardi),
(2) false memories (Deese),
(3) subjective organization (Tulving, Zaromb),
(4) working memory and attention (Jefferies),
(5) iterated learning (Kirby)
}

\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}

### 2016-05-30 brains-protocol.tex
chore/todos: update some todos from Gureckis' edited version

@@

We used a quotation dataset collected by \citet{Leskovec09}, large enough to lend itself to statistical analysis.
This dataset consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time (from August 2008 to April 2009~---~\citealp{Leskovec09-url}).\footnote{Unfortunately, the original article~\citep{Leskovec09} does not provide additional details on the source selection methodology.}
Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person. For instance,
\begin{quote}
The Bank of England said, ``these operations are designed to address funding pressures over quarter-end.''
\end{quote}
Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words). We find for example the following variation of the above quote: \begin{quote}``these operations are \textbf{intended} to address funding pressures over quarter-end.''\end{quote}
A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see][for more details]{Leskovec09}.
This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.

@@

\medskip
%, we thus considered more recently studied variables based on semantic network properties.
We relied on the ``free association'' (FA) norms collected by \citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations. FA norms record the words that come to mind when someone is presented with a given cue% (that is the ``free association'' task)
. As \citet{Nelson04} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
%Following \citet{Griffiths07}, we first consider the directed weighted network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
% \rk{we're in fact using the unweighed version of the network. Why?}
%Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.
Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where nodes are words and edges are directed from cue to target word whenever the considered target word was produced in response to the considered cue word.
This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyze is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.

@@

\todo{\#17: explain all models and the work involved, with graphs in annex}

@@

>>>>

\todo{\#20: check text/flow/definitions for clarity against Gureckis' edited pdf}

@@

We used a quotation dataset collected by \citet{Leskovec09}, large enough to lend itself to statistical analysis.
This dataset consists of the daily crawling of news stories and blog posts from around a million online sources, with an approximate publication rate of 900k texts per day, over a nine-month period of time (from August 2008 to April 2009~---~\citealp{Leskovec09-url}).\footnote{Unfortunately, the original article~\citep{Leskovec09} does not provide additional details on the source selection methodology.}
Quotations were then automatically extracted from this corpus: each quotation is a more or less faithful excerpt of an utterance (oral or written) by the quoted person. For instance,
\begin{quote}
The Bank of England said, ``these operations are designed to address funding pressures over quarter-end.''
\end{quote}

\todo{\#20: make clear we didn't do this, it's Leskovec et al.}

Quotations were then gathered in a graph and connected according to their similarity: either because they differ by very few words (in that case, no more than one word) or because they share a certain sequence of words (in that case, at least ten consecutive words). We find for example the following variation of the above quote: \begin{quote}``these operations are \textbf{intended} to address funding pressures over quarter-end.''\end{quote}
A community detection algorithm was applied to that quotation graph to detect aggregates of tightly connected, that is sufficiently similar, groups of quotations~\citep[see][for more details]{Leskovec09}.
This analysis yielded the final data we had access to, with a total of about \num{70000} sets of quotations; each of these sets allegedly contains all variations of a same parent utterance, along with their respective publication URLs and timestamps.

@@

\todo{\#20: FA don't encode reformulation explicitly, they did norms. We hypothesize it's what's at work.}

%, we thus considered more recently studied variables based on semantic network properties.
We relied on the ``free association'' (FA) norms collected by \citet{Nelson04} which naturally embed information on the idea association process underlying transformation of quotations. FA norms record the words that come to mind when someone is presented with a given cue% (that is the ``free association'' task)
. As \citet{Nelson04} explain, ``free association response probabilities index the likelihood that one word can cue another word to come to mind with minimal contextual constraints in effect.''
%Following \citet{Griffiths07}, we first consider the directed weighted network formed by the association norms, that is the network where words are nodes and edges are directed from cue to associated word, with a weight equal to the probability of that target word being produced when this particular cue was presented.
% \rk{we're in fact using the unweighed version of the network. Why?}
%Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where words are nodes and edges are directed from cue to target word whenever a target word is being produced when this particular cue word was presented.
Following \citet{Griffiths07}, we first build a directed unweighted network based on association norms, where nodes are words and edges are directed from cue to target word whenever the considered target word was produced in response to the considered cue word.
This network is of particular interest since it measures the \emph{in-vitro forced-choice} version of a substitution whereas the data we analyze is the \emph{in-vivo spontaneous} version of what we otherwise hypothesize to be the same process.

@@

\todo{\#17: explain all models and the work involved, with graphs and explanatory figures for each model in annex}

@@

\todo{\#16: explicit all significance claims, referring to the CIs}
